### 2016-07-22 code/combine_years.Rmd
Added compressed data files

---
title: "finalproject"
author: "William Paul"
date: "July 14, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(readr)
library(ggplot2)
library(dplyr)
files96 = sapply(paste("hads", c(seq(1985, 2001, by = 2)), ".txt", sep =""), read_csv)
files96 = sapply(files96, function(x){mutate(x, L30 = 0.6 * L50, GL30 = 0.6 * GL50, ABL30 = 0.6 * ABL50)})
files03 = paste("hads", c(seq(2003, 2013, by = 2)), ".txt", sep ="")
files03 = sapply(files, read_csv)

```