### 2016-07-22 code/combine_years.Rmd
Added correct compression, and initial code to combine datasets on modern ones

---
title: "finalproject"
author: "William Paul"
date: "July 14, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(readr)
library(dplyr)
mutateyears = function(data, year) {
  mutate(data, dataset_year=year)
}
files96 = lapply(paste("hads", c(seq(1985, 2001, by = 2)), ".bz2", sep =""), read_csv)
files96 = lapply(files96, data.frame, stringsAsFactors= F)
files96 = lapply(files96, function(x){mutate(x, L30 = 0.6 * L50, GL30 = 0.6 * GL50, ABL30 = 0.6 * ABL50)})
files96 = mapply(mutateyears, files03, as.list(seq(1985,2001,by=2)))
files03 = paste("hads", c(seq(2003, 2013, by = 2)), ".bz2", sep ="")
files03 = lapply(lapply(files03, read_csv), data.frame, stringsAsFactors= F)
files03 = mapply(mutateyears, files03, as.list(seq(2003,2013,by=2)))
new2011 = files03[[4]] %>% select(-VCHRMOV)
c = rbind(do.call(rbind, files03[c(1,2,3,5,6)]), new2011)
```