### 2016-08-08 presentation/index.Rmd
fixes

---
title       : Housing Affordability Over the Years
subtitle    : An Exploration into America's Housing Sector
author      : Nikolai, William, Sri, Lamb
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

```{r, echo = F, message = F}

library(ggplot2)
library(dplyr)
library(readr)
library(stringr)
library(xml2)
library(rvest)
library(tidyr)
library(purrr)
library(scales)
library(maps)
library(ggthemes)

```

```{r, echo = F}

clean_years <- read_csv("../clean_data/clean_years.csv")
regionNames = c("1" = "Northeast", "2" = "Midwest", "3" = "South", "4" = "West")

```

## Introduction: Motivation and Dataset

* Our data extends from

* William's cleaning
* Sri's cleaning
* Plots
* Conclusion stuff

---

## Cleaning the Data

Combined Years:

- Dataset is split into fifteen datasets, one for each odd year between 1985 and 2013
- To standardize the data used data from 2003 - 2013
- Reorder the columns to exactly match that of our base (2003 onward)
- Combine every list of datasets into single dataset

Clean Data:

- Initally removed NAs, expressed column in read_csv
- Set column names to lowercase, removed quotations
- Created column for owning and renting data (1= own, 2= rent)
- Write to csv (same for combined years)

---

## Variables According to the HUD Document Data Set

Variables  | Meanings
---------- | --------
Burden | Household’s monthly housing cost divided by its monthly income
Per | Household Size 
ABLMED | Median Income Adjusted for # of Bedrooms
APLMED | Median Income Adjusted for # of Persons
ZSMHC | Monthly housing costs
ZINC2 | Household Income
Assisted | Asssisted Housing
Year | The Year of the Data
Region | Census Region
fmtincrelpovcat |HH Income Relative to Poverty Income 
fmtstructuretype | Structure Type
increlamipct | HH Income relative to AMI (Percent)
fmtcostmedrelamicat | CostMed Relative to Median Income

---

## Household Income and Burden

```{r, echo = F, fig.align = 'center', fig.width = 12, message = F}

clean_years %>% 
  select(zinc2, year, region, fmtburden)%>% 
  filter(fmtburden != ".", fmtburden != "4 No Income") %>% 
  mutate(fmtburden = str_replace(fmtburden, 
                                 "Greater than 50%", 
                                 "50% or More")) %>% 
  group_by_("year", "fmtburden") %>%  
  summarise_(new_y = as.formula(
  paste0("~ median(","zinc2", ",na.rm=T)"))) %>%

  ggplot(aes_string(x = "year", y = "new_y", color = "fmtburden")) + 
  geom_line(size = 1) + 
  labs(title = "Relationship between Household \nIncome and Burden", 
       y = "Household Income", x = "Year", color = "Burden") +
  scale_y_continuous(limits=c(0,NA),
                     breaks=c(0,20000,40000,60000),
                     labels=function(x){paste0(x/1000, "K")}) +
  scale_colour_brewer(palette = "Dark2") +
  theme_tufte(base_size = 20)

```

---

## Monthly Housing Costs and Burden

```{r, echo = F, fig.align = 'center', fig.width = 12}

cleanyearsmonthly <- clean_years %>% 
  select(year, zsmhc, region, fmtburden) %>% 
  filter(fmtburden != ".") %>% 
  mutate(fmtburden = str_replace(fmtburden, "Greater than 50%", "50% or More"))

####overall graph
cleanyearsmonthly %>%  
  group_by_("year", "fmtburden") %>%  
  summarise_(new_y = as.formula(
    paste0("~ median(","zsmhc", ",na.rm=T)"))) %>%

  ggplot(aes_string(x = "year", y = "new_y", color = "fmtburden")) + 
  geom_line(size = 1) + 
  labs(title = "Relationship between Monthly \nHousing Costs and Burden", 
       y = "Monthly Housing Costs ($)", x = "Year", color = "Burden") +
  scale_colour_brewer(palette = "Dark2") +
  theme_tufte(base_size = 20)

```

---

## Regional Monthly Housing Costs and Burden

```{r, echo = F, fig.align = 'center', fig.width = 12, message = F }

  cleanyearsmonthly %>%  
  group_by_("year","region", "fmtburden") %>%  
  summarise_(new_y = as.formula(
    paste0("~ median(","zsmhc", ",na.rm=T)"))) %>% 
  ggplot(aes_string(x = "year", y = "new_y", color = "fmtburden")) + 
  geom_line(size = 1) + 
  facet_wrap(~region, labeller = as_labeller(regionNames)) + 
  labs(title = "Relationship between Monthly \nHousing Costs and Burden", 
       y = "Monthly Housing Costs ($)", x = "Year", color = "Burden") +
  scale_colour_brewer(palette = "Dark2") +
  theme_tufte(base_size = 20)

```

---

## Median Income Among Differing Census Regions in 2013

```{r, echo = F, fig.align = 'center', fig.width = 12, message = F}

url <- "http://www.mapsofworld.com/usa/usa-maps/united-states-regional-maps.html"
html <- read_html(url)
xpath <- '//table[@class="tableizer-table"]'

regions_tbl <- html %>% 
  html_nodes(xpath = xpath) %>%
  html_table %>%
  .[[1]]

region_1 <- regions_tbl[2:7,] %>% 
  gather(key = region, value = state, 1:2) %>% 
  slice(1:9) %>%
  mutate(region = 1)

region_2 <- regions_tbl[10:15,] %>% 
  gather(key = region, value = state, 1:2) %>% 
  mutate(region = 2)

# Add missing Missouri
region_2[6, 2] <- "Missouri"

region_3 <- regions_tbl[18:27,] %>%  
  gather(key = region, value = state, 1:2) %>% 
  slice(c(-10, -15, -16)) %>%
  mutate(region = 3)

region_4 <- 
  regions_tbl[30:37,] %>%  
  gather(key = region, value = state, 1:2) %>% 
  slice(1:13) %>%
  mutate(region = 4)

regions <- rbind(region_1, region_2, region_3, region_4) %>%
  mutate(state = str_to_title(state),
         region = as.character(region))

median_incomes <- clean_years %>%
  ##filter(year == 2013) %>%
  mutate(region = as.character(region)) %>%
  group_by(region) %>%
  summarise(zinc2_median = median(zinc2, na.rm = T))

map_df <- map_data("state") %>%
  mutate(region = str_to_title(region)) %>%
  rename(state = region)  %>%
  # Join with census region information
  left_join(regions) %>%
  left_join(median_incomes)

map_df %>% 
  ggplot() +  
  geom_polygon(aes(x=long, 
                   y = lat, 
                   fill = zinc2_median, 
                   group = group)) + 
  coord_fixed(1.3) +
  theme_tufte(base_size = 20) +
  scale_x_continuous(name = NULL,
                     labels = NULL,
                     breaks = NULL) +
  scale_y_continuous(name = NULL,
                     labels = NULL,
                     breaks = NULL) +
  labs(fill = "Median income") +
  scale_fill_distiller(limits = c(25000, 42000), 
                                 breaks = sort(unique(map_df$zinc2_median),
                                               decreasing = T),
                                 palette = "Greens", 
                                 direction = 1,
                                 guide = "legend",
                        labels=function(x){paste0("$",x %/% 100 / 10, "K")})

```

---

## Household Size and Burden

```{r, echo = F}

  clean_years %>%
    mutate(per = cut(per, 
                     breaks = c(0, 1, 2, Inf),
                     labels = c("1", "2", "3 or more")))  %>%
    group_by(year, per) %>%  
    # Take the median instead of mean to avoid extreme observations skewing the 
    # result.
    summarise(new_y = median(burden, na.rm = T)) %>%
    ggplot(aes(x = year,   
                      y = new_y,   
                      colour = per)) +  
    geom_line(size = 1) +
    labs(colour = "Household Size",
         x = "Year",
         y = "Median Burden",
         title = "Relationship between Burden and Household Size") +
    scale_colour_brewer(palette = "Dark2") +
  scale_y_continuous(labels=function(x){paste0(x*100, "%")}) +
    theme_tufte(base_size = 20)

```

---

## Households and Poverty

```{r, echo = F, fig.align = 'center', fig.width = 12}
clean_years %>% 
  # Group and summarise the amount of people in each poverty group each year
  group_by(fmtincrelpovcat, year) %>%
  summarise(count = n()) %>%
  # Clean NAs
  na.omit() %>%
  # Group again by year to calculate relative percentages per year
  group_by(year) %>% 
  mutate(percent = count / sum(count),  
         # Clean the labels
         fmtincrelpovcat = factor(fmtincrelpovcat, 
                                  levels = c("1 LTE Poverty", 
                                             "2 100-150% Poverty", 
                                             "3 150-200% Poverty", 
                                             "4 200%+ Poverty"),  
                                  labels = c("Less than poverty income",   
                                             "100-150% of poverty income",   
                                             "150-200% of poverty income",    
                                             "> 200% of poverty income"),
                                  ordered = T)) %>% 
  ggplot(aes(x = year, 
             y = percent,
             fill = fmtincrelpovcat)) + 
  geom_area(alpha = 0.8) +
  # Use a nice colour palette
  scale_fill_brewer(palette = "RdYlGn", 
                    # Reverse the order of legend items to match the
                    # stacking order
                    guide = guide_legend(reverse=TRUE)) +
  labs(title = "Number of Households Relative to the Poverty Line",
       x = "Year",
       y = "Percent of Households",
       fill = "Household Income \nRelative to Poverty Income") +
  scale_y_continuous(labels=function(x){paste0(x*100, "%")}) +
  scale_colour_brewer(palette = "Dark2") +
  theme_tufte(base_size = 20)
```

---

## Income Compared to Area's Median by House TypeS

```{r, echo = F, fig.align = 'center', fig.width = 12, message =F}

clean_years %>% filter(fmtstructuretype != ".") %>% 
      # Use NSE
      group_by_("year", "fmtstructuretype") %>%  
      # Take the median instead of mean to avoid extreme observations skewing the 
      # result.
      summarise_(new_y= as.formula(paste0("~ median(","increlamipct", ",na.rm=T)"))) %>%  
      # Use NSE
      ggplot(aes_string(x = "year",   
                        y = "new_y",   
                        colour = "fmtstructuretype")) +  
  geom_line(size = 1) + 
  labs(colour = "Structure Type",
       y="Income percentage of area median income",
       x = "Year") +
  scale_colour_brewer(palette = "Dark2") +
  theme_tufte(base_size = 20)

```

---

## Burden Based on Median Housing Cost with Assistance

```{r, echo = F, fig.align = 'center', fig.width = 12, message= F, warning=F}

assisted = clean_years %>% filter(!is.na(assisted)) %>% 
      # Use NSE
      group_by_("assisted", "fmtcostmedrelamicat") %>%  
      # Take the median instead of mean to avoid extreme observations skewing the 
      # result.
      summarise_(new_y= as.formula(paste0("~ median(","burden", ",na.rm=T)"))) %>% 
      ungroup %>% 
      mutate(assisted = as.factor(assisted))
ggplot(assisted, aes(y = new_y,   
                  x = fmtcostmedrelamicat,
                  fill = assisted, group = assisted)) +  
  geom_bar(
           stat="identity") + 
  labs(y="Burden",
       x = "",
       fill="Assistance") + 
  scale_y_continuous(limits=c(0,1)) +
  scale_fill_brewer("Assistance", 
                    labels = c("Without Assistance", 
                               "With Assistance"),
                    palette = "Set2")  +
  coord_flip() +
  theme_tufte(base_size = 20)

```

---

## MORE GRAPHS

```{r, echo = F, fig.align = 'center', fig.width = 12, message= F}

clean_years %>%
  mutate(region = as.character(region)) %>% 
  group_by(year, region) %>%  
  # Take the median instead of mean to avoid extreme observations skewing the 
  # result.
  summarise(new_y = median(aplmed, na.rm = T)) %>%
  ggplot(aes(x = year,   
                  y = new_y,   
                  colour = region)) + 
  geom_line(size = 1) + 
  labs(x = "Year", y = "Median Adjusted Income (per persons)", 
       title = "Median Income Adjusted \nfor the Number of People by Region") + 
  scale_colour_brewer(palette = "Dark2", 
                      name = "Region", 
                      labels = regionNames) +
  theme_tufte(base_size = 20)
```

- Adjusted income per persons
- Bracketed by HUD indexes (0.70 for 1 person, 0.80 for 2, 1.00 for 4)
- Spike around 2010 associated with financial crisis

---

## Median Income Adjusted for Number of Bedrooms By Region

```{r, echo = F, fig.align = 'center', fig.width = 12, message = F}

clean_years %>%
  mutate(region = as.character(region)) %>% 
  group_by(year, region) %>%  
  # Take the median instead of mean to avoid extreme observations skewing the 
  # result.
  summarise(new_y = median(ablmed, na.rm = T)) %>%
  ggplot(aes(x = year,   
                  y = new_y,   
                  colour = region)) + 
  geom_line(size = 1) +
  labs(x = "Year", y = "Median Adjusted Income (Per Bedroom)") + 
  scale_colour_brewer(palette = "Dark2",
                      name = "Region", 
                      labels = regionNames) +
  theme_tufte(base_size = 20)

```

Adjusted income per bedroom

Bracketedby HUD indexes (0.70 for 0, 0.90 for 2, 1.04 for 3)

More discrepancy in burden levels between the Midwest/South and Northeast/West

  - Larger homes in the Midwest/South

Purhasing parity shoudl be higher in midwest/south, not expressed in data

---

## Concluding Thoughts

Housing Affordability Has Decreased

The 2007 Housing Bubble Crisis and it's Consequences

Census Regions Differences

  - Western and Northeast Regions Vs. South and Midwest

---

## Future Questions

* Does the Data in some graphs keep showing a downward slope?

* If we analyzed data from 2013 - 2016, would it show another potential, housing housing bubble?

* Our data is limited to 2013, so what is the housing situation most recently?

---

## Ask Us Anything

---